---
title: "Aggregation Example"
author: 
  - Rastin Matin
  - Benjamin Christoffersen 
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:  
  html_document:
    toc: true
bibliography: refs.bib
---

<script>
$(document).ready(function(){
  var hide_divs = $("div.hideable");
  hide_divs.each(function(){
    // Wrap content in div
    $(this).wrapInner( "<div class='hideable_content', style='display: none;'></div>");
    
    // Add button
    $(this).prepend("<button id='toogle'>show</button>");
  });
  
  // Add hideable btn
  // Put the rest in a div
  
  $("div.hideable button#toogle").click(function(){
    var parent = $(this).parent();
    var target_div = $(parent).find("div.hideable_content");
    
    if(target_div.css("display") == "none"){
      target_div.show();
        $(this).text("Hide");
    } else {
      target_div.hide();
      $(this).text("Show");
    }
  });
});
</script>

## Load data

```{r static_setup, include=FALSE, cache=FALSE}
# please do not set options here that could change...
knitr::opts_chunk$set(
  cache.path = 
    paste0(file.path("cache", "aggregation-example"), .Platform$file.sep), 
  fig.path = 
    paste0(file.path("fig"  , "aggregation-example"), .Platform$file.sep))
```

```{r setup, include=FALSE, cache=FALSE}
# please do set options here that could change...
knitr::opts_chunk$set(
  echo = TRUE, fig.height = 4, fig.width = 7, dpi = 72, comment = "#R", 
  error = FALSE)
options(digits = 4, scipen = 10, width = 90)
```

```{r run_setup, child = 'setup.Rmd'}
```

## Show Densities Through Time

```{r load_gg}
library(ggridges)
library(ggplot2)
library(parallel)
```

```{r sigma_do_dens, cache = 1}
local({
  dat$time_var <- format(make_ym_inv(dat$tstop), "%Y")
  dat$siga <- c(wz(dat$sigma, do_center = FALSE, scale = FALSE))
  
  get_plot_device({
    ggplot(subset(dat, as.integer(time_var) > 1998L), 
           aes(x = siga, y = time_var)) + 
      stat_density_ridges(scale = 3, fill = rgb(0, 0, 0, 0), quantile_lines = TRUE, 
                          quantiles = 2) + theme_ridges() +
      scale_x_continuous(expand = c(0.01, 0)) +
      scale_y_discrete(expand = c(0.01, 0)) + 
      labs(x = get_label("sigma"), y = "Year (t - 1, t]")
  }, file_name_wo_extension = "agg-sigma-time-dens", ggplot = TRUE)
})
```

## Example in Paper

```{r def_sim_func}
#####
# assign function to simulate outcome
n_periods <- 64L + 1L

# means for mixture component with time-varying means
mea_min <- -3
mea_max <- 2
mid <- (n_periods - 1L) / 2L 
tfunc <- abs(1:n_periods - mid - 1L)
tfunc <- (tfunc - min(tfunc)) / (max(tfunc) - min(tfunc)) * 
  (mea_max - mea_min) + mea_min
plot(tfunc, type = "l")

# assign simulation function
get_dat <- function(){
  sim_dat <- replicate(1000, {
    # draw starting time
    tstart <- (sample.int(n_periods, 1L) - 1L):(n_periods - 1L)
    
    # draw covariates
    n <- length(tstart)
    X <- ifelse(
      .2 > runif(n), 
      rnorm(n * 1L, mean = tfunc[tstart + 1L], sd = .5), 
      rnorm(n * 1L))
    
    # determine if individual has an event and remove observations after
    # if so
    y <- -expm1(-exp(X - 3)) > runif(n) 
    idx <- if(any(y)) 1:min(which(y)) else 1:n
    
    # return 
    data.frame(X, y = y, tstop = tstart + 1L, tstart = tstart)[idx, ]
  }, simplify = FALSE)
  do.call(rbind, sim_dat)
}

# show example
set.seed(18404064)
tmp <- get_dat()
ggplot(tmp, aes(x = X, y = as.factor(tstop))) + 
  stat_density_ridges(scale = 3, fill = rgb(0, 0, 0, 0), quantile_lines = TRUE, 
                      quantiles = 2) + theme_ridges() +
  scale_x_continuous(expand = c(0.01, 0)) +
  scale_y_discrete(expand = c(0.01, 0)) + labs(y = "tstop")
```

```{r run_sim, cache = 1}
######
# simulate

# knot positions
(knt <- seq(1, n_periods, length.out = 5L)[-c(1, 5)])

# simulate and compute
cl <- makeCluster(6L)
set.seed(28789699)
clusterExport(cl, c("knt", "n_periods", "n_periods", "get_dat", "tfunc"), 
              envir = environment())
clusterSetRNGStream(cl)

out <- parLapply(cl, 1:1000L, function(...){
  library(splines)
  
  # simulate data and fit true model
  sim_dat <- get_dat()
  f1 <- glm(y ~ X, binomial("cloglog"), sim_dat)
  
  # aggregate data 
  sim_col <- lapply(split(sim_dat, sim_dat$tstop), function(x){
    n <- nrow(x)
    with(x, data.frame(n = n, y = sum(y) / n, X = mean(X), tstart = tstart[1L], 
                       tstop = tstop[1L]))
  })
  sim_col <- do.call(rbind, sim_col)
  
  # fit models to aggregate data
  f2 <- glm(y ~ X, binomial("cloglog"), sim_col, weights = n)
  
  # add splines
  f3 <- update(f2, . ~ . + ns(tstop, knots = knt))
  
  list(f1 = f1, f2 = f2, f3 = f3)
}, simplify = FALSE)

stopCluster(cl)
```

```{r gather_res}
#####
# gather result
ests <- local({
  ests <- lapply(out, function(o) sapply(o, function(m) coef(m)))
  out <- lapply(seq_along(ests[[1L]]), function(i) sapply(ests, "[[", i))
  names(out) <- names(ests[[1L]])
  out
})

# MC means and standard errors
(summas <- lapply(ests, function(x) apply(x, 1L, function(z){
  mea <- mean(z)
  sdo <- sd(z) / sqrt(length(z))
  c(mean = mea, sd = sdo)
})))

# likelihood ratio tests
z <- lapply(out, function(x) anova(x$f2, x$f3, test = "LRT"))
ps <- sapply(lapply(z, "[", TRUE, "Pr(>Chi)"), "[[", 2L)
plot(ecdf(ps))
abline(a = 0, b = 1, lty = 2)
sum(ps < 0.05)
quantile(ps, seq(0, 1, .01))
```
